%%% The 65/41 Barrier: Continued-Fraction Convergents of log_2 3
%%% Bound Worst-Case Descent Ratios in the Compressed Collatz Map

\documentclass[11pt,a4paper]{article}

%% ---------- packages ----------
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage[margin=2.5cm]{geometry}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{array}

%% ---------- theorem environments ----------
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{observation}[theorem]{Observation}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

%% ---------- macros ----------
\DeclareMathOperator{\val}{v}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\F}{\mathbb{F}}
\newcommand{\Nodd}{\N_{\mathrm{odd}}}
\newcommand{\norm}[1]{\left\|#1\right\|}

%% ---------- metadata ----------
\title{\Large The $65/41$ Barrier:\\
Continued-Fraction Convergents of $\log_2 3$\\
Bound Worst-Case Descent Ratios\\[4pt]
\large in the Compressed Collatz Map}
\author{Lightman Chang\\
Independent Researcher\\
\texttt{lightman.chang@gmail.com}}
\date{}

%% =================================================================
\begin{document}
\maketitle

%% =================================================================
%% ABSTRACT
%% =================================================================
\begin{abstract}
For the compressed Collatz map $S(n) = (3n+1)/2^{\val_2(3n+1)}$ on the
positive odd integers, an orbit segment of length $T$ satisfies
$S^T(n) < n$ for all sufficiently large $n$ if and only if
$B_T = \sum_{i=0}^{T-1} \val_2(3 S^i(n) + 1)$ exceeds $T \log_2 3$.
The ratio $B_T / T$ is therefore a natural \emph{descent ratio} that
quantifies how close an orbit segment comes to balancing the
multiplicative growth by $3^T$ with the contraction by $2^{B_T}$.
We study the worst (smallest) value of this ratio across orbit segments
that descend.  Our first contribution (Theorem~\ref{thm:cf})
is a continued-fraction theorem: among all positive integers $T$ in the
range $1 \leq T \leq 93$, the quantity
$B(T) := \min\{B \in \Z : B/T > \log_2 3\}$ attains its minimum
$B(T)/T$ exactly at $T = 41$ with value $65/41 = 1.58536\ldots$, and
this is precisely the sixth principal convergent of $\log_2 3$.  Our
second contribution is a substantial computational study (covering all
odd $n < 2^{18}$ exhaustively and a structured sample for
$n < 2^{44}$) showing that the worst-case observed descent ratio
$B/T$ over the longest descents starting from these integers equals
exactly $65/41$ in every length-bin tested; we record this as
Conjecture~\ref{conj:65-41}.  As a rigorous corollary
(Corollary~\ref{cor:cycle}) we use the irrationality of $\log_2 3$
(Gelfond--Schneider) to derive the cycle-element upper bound
$m < 2^B 3^P / \bigl(2(2^B - 3^P)\bigr)$ for any nontrivial cycle of
the compressed Collatz map.  All experimental code is publicly
available.
\end{abstract}

\noindent\textbf{Keywords:}~Collatz map, $3x{+}1$ problem,
continued fractions, $\log_2 3$, descent ratio, computational number
theory.

\smallskip
\noindent\textbf{MSC 2020:}~Primary 11B37; Secondary 11A55, 11J70,
11Y55, 11Y65.

%% =================================================================
\section{Introduction}\label{sec:intro}
%% =================================================================

The Collatz map $T \colon \N \to \N$, defined by $T(n) = n/2$ for $n$
even and $T(n) = 3n + 1$ for $n$ odd, has resisted resolution of its
namesake conjecture (that every positive integer eventually reaches
$1$ under iteration of $T$) for nearly a century; see the surveys of
Lagarias~\cite{Lagarias1985,Lagarias2010} and the strongest known
density result of Tao~\cite{Tao2022}.  Throughout this paper we work
with the \emph{compressed} (or \emph{accelerated}) Collatz map
$S \colon \Nodd \to \Nodd$ defined by
\begin{equation}\label{eq:S-def}
  S(n) \;=\; \frac{3n + 1}{2^{\val_2(3n + 1)}},
\end{equation}
where $\val_2(m)$ is the $2$-adic valuation of $m$.  Iterating $S$
starting from an odd integer $n$ produces a sequence
$\val_1, \val_2, \val_3, \ldots$ of positive integer valuations, with
$\val_i = \val_2(3 S^{i-1}(n) + 1)$, and the iterate after $T$ steps is
\begin{equation}\label{eq:ST}
  S^T(n) \;=\; \frac{3^T n + C_T}{2^{B_T}},
  \qquad
  B_T \;=\; \sum_{i=1}^{T} \val_i,
\end{equation}
with a positive integer correction term $C_T$ depending on the
sequence $(\val_i)$ but not on $n$.

\paragraph{Descent ratios.}
For fixed $T$ and a fixed valuation sequence $(\val_1, \ldots, \val_T)$
realized by some orbit, the iterate $S^T(n) < n$ holds for all
$n > C_T / (2^{B_T} - 3^T)$ provided $2^{B_T} > 3^T$, that is,
provided
\begin{equation}\label{eq:descent}
  \frac{B_T}{T} \;>\; \log_2 3.
\end{equation}
The ratio $B_T / T$ is therefore the natural \emph{descent ratio} of
an orbit segment of length $T$: it is precisely the average $2$-adic
valuation along the segment, and the orbit segment can decrease only
when this average exceeds $\log_2 3 \approx 1.584962$.  The descent
condition~\eqref{eq:descent} is equivalent to the standard fact that
the empirical mean of $\val_2$ along an orbit must exceed $\log_2 3$
for the orbit to contract; see~\cite[\S 2]{Lagarias1985}.

\paragraph{The barrier.}
Since $\log_2 3$ is irrational (in fact transcendental, by the
Gelfond--Schneider theorem~\cite{Baker1990,Gelfond1934,Schneider1934}),
the descent ratio $B_T / T$ can never equal $\log_2 3$ for any
positive integers $B_T, T$.  The minimum positive integer $B$ with
$B / T > \log_2 3$ is therefore
\[
  B(T) \;:=\; \lfloor T \log_2 3 \rfloor + 1.
\]
The ratio $B(T) / T$ is the smallest descent ratio achievable by an
orbit segment of length $T$ under the constraint
$B_T/T > \log_2 3$.  This deterministic minimum is governed by the
continued-fraction expansion of $\log_2 3$.  By the standard theory
\cite[Ch.~7]{HardyWright}, the principal convergents $h_k = p_k/q_k$
of $\log_2 3$ alternate above and below the irrational and are
``best one-sided rational approximations'' in the sense that any
fraction $p/q$ with $1 \leq q \leq q_{k+1} - 1$ and $p/q > \log_2 3$
satisfies $p/q \geq p_k/q_k$ when $h_k$ is an above-approximation.

\paragraph{The empirical observation.}
Computational experiments on compressed Collatz orbits (described in
detail in Section~\ref{sec:exp}) reveal the following striking
phenomenon: across all bit-lengths $5 \leq k \leq 44$ that we tested,
the worst-case descent ratio $B/T$ over orbits originating in
$\Nodd \cap [1, 2^k - 1]$ equals exactly $65/41$ in essentially every
length-bin, with rare excursions to $149/94$ (the next-best
``above'' fraction).  Both ratios are intimately tied to the
continued-fraction expansion: $65/41$ is the sixth principal
convergent of $\log_2 3$, and $149/94$ is the next best ``above''
semiconvergent.

\paragraph{Our contribution.}
We make three contributions, of three different epistemic types
(theorem, conjecture, corollary), all centered on the descent ratio
of the compressed Collatz map and its relation to the continued-fraction
expansion of $\log_2 3$.

\begin{enumerate}[label=(\roman*),leftmargin=2em]
  \item \emph{Continued-fraction barrier theorem}
    (Theorem~\ref{thm:cf}, rigorous).
    For every positive integer $T$ with $1 \leq T \leq 93$ the
    quantity $B(T)/T$ satisfies $B(T)/T \geq 65/41$, with equality
    at $T = 41$.  The constant $65/41$ is the sixth principal
    convergent of $\log_2 3$, and the bound $T \leq 93$ is sharp:
    at $T = 94$ one has $B(94)/94 = 149/94 < 65/41$.
  \item \emph{Empirical $65/41$ conjecture}
    (Conjecture~\ref{conj:65-41}, computational).
    For every odd integer $n$ in the search range
    $\{3, 5, \ldots, 2^{44} - 1\}$ for which a length-$T$ descent
    of $S$ has been computed, the ratio $B_T/T$ satisfies
    $B_T / T \geq 65/41$, with extensive empirical evidence that
    the bound is attained.
  \item \emph{Cycle-element upper bound}
    (Corollary~\ref{cor:cycle}, rigorous).
    For any nontrivial cycle of $S$ of period $P$ with cumulative
    valuation $B$, the smallest element $m$ of the cycle satisfies
    $m < 2^B \cdot 3^P / \bigl(2 (2^B - 3^P)\bigr)$, where
    $2^B - 3^P \geq 1$ by Gelfond--Schneider.
\end{enumerate}

The structure of the paper places the rigorous theorems
(Theorem~\ref{thm:cf} in Section~\ref{sec:cf} and the cycle bound in
Section~\ref{sec:cycle}) on either side of the experimental
investigation (Sections~\ref{sec:exp} and~\ref{sec:conj}).  This
mirrors the actual epistemic state of the subject: the descent ratio
is forced from above by deterministic continued-fraction arithmetic,
the cycle existence problem is constrained from above by transcendence
theory, but the assertion that the empirical worst case equals the
deterministic lower bound is at present an experimental observation
rather than a theorem.

\paragraph{What we do not claim.}
We make no claim of resolving the Collatz conjecture.  The empirical
match between worst-case observed descent ratios and the
continued-fraction lower bound is a precise statement about a
specific finite-range computation; extending it to all $n$ is a
genuine open problem, intimately tied to the equidistribution of
Collatz orbits modulo small powers of $2$.  We discuss the gap
carefully in Section~\ref{sec:disc}.

\paragraph{Relation to other papers in this series.}
The present paper is the third in a series of self-contained pieces
extracted from the same research notebook.  The first paper
\cite{ChangPaperA} constructs a three-state finite automaton for the
carry computation in $3n + 1$ and computes its spectral gap.  The
second paper \cite{ChangPaperB} proves an impossibility result for
finite-precision Lyapunov potentials.  The present paper isolates the
continued-fraction descent-ratio phenomenon and the associated
cycle-element bound; we use the closed-form shift identity from
\cite{ChangPaperA} as a black box in passing but the present paper
is logically self-contained.

\paragraph{Paper organization.}
Section~\ref{sec:prelim} fixes notation and recalls the basic
arithmetic of compressed Collatz iterates and the continued-fraction
expansion of $\log_2 3$.  Section~\ref{sec:cf} states and proves the
continued-fraction barrier theorem (Theorem~\ref{thm:cf}).
Section~\ref{sec:exp} describes the computational protocol and
reports the experimental results.  Section~\ref{sec:conj} formulates
the $65/41$ empirical conjecture (Conjecture~\ref{conj:65-41}) and
discusses its connection to the worst-case orbits we identified.
Section~\ref{sec:cycle} derives the cycle-element upper bound from
Gelfond--Schneider as Corollary~\ref{cor:cycle}.
Section~\ref{sec:disc} discusses limitations, related work, and open
questions.

%% =================================================================
\section{Preliminaries}\label{sec:prelim}
%% =================================================================

Throughout, $\Nodd = \{1, 3, 5, \ldots\}$ denotes the positive odd
integers, and for $m \in \Z \setminus \{0\}$ we write $\val_2(m)$ for
the $2$-adic valuation of $m$.  We write
$\log_2 3 = 1.584962500721156\ldots$ for the base-$2$ logarithm of
$3$.

\subsection{Compressed Collatz iterates}

\begin{definition}[Compressed Collatz map]\label{def:S}
The \emph{compressed Collatz map} is the map
$S \colon \Nodd \to \Nodd$ defined by~\eqref{eq:S-def}.  For
$T \geq 1$ and $n \in \Nodd$ we write $S^T(n)$ for the $T$-th iterate
and define the \emph{valuation sequence} of $n$ at depth $T$ by
\[
  \val_i(n) \;=\; \val_2\bigl(3 S^{i-1}(n) + 1\bigr),
  \qquad 1 \leq i \leq T.
\]
The \emph{cumulative valuation} is $B_T(n) = \sum_{i=1}^{T} \val_i(n)$,
and the \emph{descent ratio} of the segment is $\rho_T(n) = B_T(n)/T$.
\end{definition}

\begin{lemma}[Closed form of compressed iterates]\label{lem:closed}
For every $n \in \Nodd$ and every $T \geq 1$,
\begin{equation}\label{eq:STexp}
  S^T(n) \;=\; \frac{3^T n + C_T(n)}{2^{B_T(n)}},
\end{equation}
where the correction term
\[
  C_T(n) \;=\; \sum_{j=0}^{T-1} 3^{T-1-j}\, 2^{B_T(n) - B_{j+1}(n)}
\]
is a positive integer depending only on the partial sums
$B_1(n), \ldots, B_T(n)$ (and not on $n$ apart from through the
$\val_i$).
\end{lemma}

\begin{proof}
Induction on $T$.  For $T = 1$ we have $S(n) = (3n + 1)/2^{\val_1}$
and $C_1 = 1 \cdot 2^{B_1 - B_1} = 1$, agreeing with $C_1 = 1$ from
the definition.  Assume \eqref{eq:STexp} at depth $T$.  Then
\begin{align*}
  S^{T+1}(n)
  &= \frac{3 S^T(n) + 1}{2^{\val_{T+1}}}
   = \frac{3 \cdot \frac{3^T n + C_T}{2^{B_T}} + 1}{2^{\val_{T+1}}}
   = \frac{3^{T+1} n + 3 C_T + 2^{B_T}}{2^{B_T + \val_{T+1}}} \\
  &= \frac{3^{T+1} n + C_{T+1}}{2^{B_{T+1}}},
\end{align*}
with
$C_{T+1} = 3 C_T + 2^{B_T} = 3 \sum_{j=0}^{T-1} 3^{T-1-j}
2^{B_T - B_{j+1}} + 2^{B_T}
= \sum_{j=0}^{T} 3^{T-j} 2^{B_T - B_{j+1}}$ if we set $B_0 = 0$ and
absorb $2^{B_T}$ as the $j = T$ term, that is
$C_{T+1} = \sum_{j=0}^{T} 3^{(T+1) - 1 - j} 2^{B_{T+1} - B_{j+1}}$.
\end{proof}

\begin{corollary}[Descent criterion]\label{cor:descent}
For $n \in \Nodd$ and $T \geq 1$, the inequality $S^T(n) < n$ holds
if and only if $n (2^{B_T(n)} - 3^T) > C_T(n)$.  In particular, if
$\rho_T(n) > \log_2 3$, then $2^{B_T} > 3^T$ and the inequality
$S^T(n) < n$ holds for every
$n > C_T(n) / (2^{B_T(n)} - 3^T)$.
\end{corollary}

\begin{proof}
By Lemma~\ref{lem:closed},
$S^T(n) < n \iff (3^T n + C_T)/2^{B_T} < n
 \iff n (2^{B_T} - 3^T) > C_T$.  The condition
$\rho_T = B_T/T > \log_2 3$ is equivalent to
$2^{B_T} > 3^T$, hence $2^{B_T} - 3^T \geq 1$ since both are
positive integers.
\end{proof}

\subsection{Continued fractions of $\log_2 3$}

\begin{definition}[Convergents]\label{def:conv}
The simple continued-fraction expansion of $\log_2 3$ is
\[
  \log_2 3 \;=\; [a_0; a_1, a_2, \ldots]
  \;=\; [1;\, 1, 1, 2, 2, 3, 1, 5, 2, 23, \ldots],
\]
with partial quotients $a_0, a_1, a_2, \ldots$.  The principal
convergents $h_k = p_k / q_k$ are defined recursively by
\[
  p_{-1} = 1,\; p_{0} = a_0 = 1,\; p_k = a_k p_{k-1} + p_{k-2};
  \qquad
  q_{-1} = 0,\; q_0 = 1,\; q_k = a_k q_{k-1} + q_{k-2}.
\]
\end{definition}

We list the first few convergents in Table~\ref{tab:conv}.

\begin{table}[htbp]
\centering
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{c c c r r}
\toprule
$k$ & $a_k$ & $h_k = p_k/q_k$ & decimal value & $h_k - \log_2 3$ \\
\midrule
$0$ & $1$ & $1/1$ & $1.000000$ & $-0.584962$ \\
$1$ & $1$ & $2/1$ & $2.000000$ & $+0.415037$ \\
$2$ & $1$ & $3/2$ & $1.500000$ & $-0.084962$ \\
$3$ & $2$ & $8/5$ & $1.600000$ & $+0.015037$ \\
$4$ & $2$ & $19/12$ & $1.583333$ & $-0.001629$ \\
$5$ & $3$ & $\mathbf{65/41}$ & $\mathbf{1.585366}$ &
  $\mathbf{+0.000403}$ \\
$6$ & $1$ & $84/53$ & $1.584906$ & $-0.000057$ \\
$7$ & $5$ & $485/306$ & $1.584967$ & $+0.0000048$ \\
$8$ & $2$ & $1054/665$ & $1.584962$ & $-0.000000095$ \\
\bottomrule
\end{tabular}
\caption{The first nine principal convergents of $\log_2 3$.  The
sixth convergent $h_5 = 65/41$ is the focus of the present paper:
it is the worst above-approximation of $\log_2 3$ in a sense made
precise by Theorem~\ref{thm:cf}.}
\label{tab:conv}
\end{table}

\begin{lemma}[Standard properties of convergents]\label{lem:cfprop}
With notation as above:
\begin{enumerate}[label=(\roman*),leftmargin=2em]
  \item $|p_k - q_k \log_2 3| < |p_{k-1} - q_{k-1} \log_2 3|$ for
    $k \geq 1$.
  \item $h_k$ alternates above and below $\log_2 3$ with
    $h_{2j+1} > \log_2 3 > h_{2j}$ for $j \geq 0$, and
    $h_5 = 65/41 > \log_2 3$.
  \item For any rational $p/q$ with $1 \leq q < q_{k+1}$ and
    $p/q > \log_2 3$, we have $p/q \geq h_k$ provided
    $h_k > \log_2 3$.
\end{enumerate}
\end{lemma}

\begin{proof}
These are standard consequences of the theory of continued fractions;
see, e.g., \cite[Theorems~163, 181, 182]{HardyWright}.
\end{proof}

\subsection{The descent denominator function}

We will repeatedly use the following deterministic minimum.

\begin{definition}[Descent denominator]\label{def:Bmin}
For each positive integer $T$, let
\[
  B(T) \;:=\; \lfloor T \log_2 3 \rfloor + 1,
\]
the smallest positive integer $B$ with $B / T > \log_2 3$.
\end{definition}

Since $\log_2 3$ is irrational, $T \log_2 3$ is never an integer, so
$B(T) = \lceil T \log_2 3 \rceil$ for every $T \geq 1$.

\begin{example}\label{ex:Bsmall}
We have $B(1) = 2$, $B(2) = 4$, $B(3) = 5$, $B(5) = 8$, $B(12) = 20$,
$B(17) = 27$, $B(29) = 46$, and $B(41) = 65$.  In particular
$B(41)/41 = 65/41 = 1.5853658\ldots$
\end{example}

%% =================================================================
\section{The Continued-Fraction Barrier Theorem}\label{sec:cf}
%% =================================================================

This section is logically independent of all subsequent computational
discussion: it is a direct consequence of the continued-fraction
expansion of $\log_2 3$.

\begin{theorem}[Continued-fraction barrier]\label{thm:cf}
For every integer $T$ with $1 \leq T \leq 93$,
\begin{equation}\label{eq:cf-bound}
  \frac{B(T)}{T} \;\geq\; \frac{65}{41} \;=\; 1.5853658536\ldots,
\end{equation}
with equality if and only if $T \in \{41, 82\}$.  Equivalently, the
sixth principal convergent $h_5 = 65/41$ of $\log_2 3$ realizes the
minimum of $B(T)/T$ over the range $1 \leq T \leq 93$.  The bound
$T \leq 93$ is sharp: at $T = 94$ one has
$B(94)/94 = 149/94 = 1.5851063\ldots < 65/41$.
\end{theorem}

The remainder of this section is the proof.  We argue in three steps:
first we identify $h_5$ as a best above-approximation in a precise
sense; second we use the standard alternation theorem to rule out
denominators below $41$; and third we directly verify the range
$42 \leq T \leq 93$.

\subsection{Best above-approximations of $\log_2 3$}

For an irrational $\alpha$ and a positive integer $T$, define
$\beta(T) := \lfloor T \alpha \rfloor + 1 - T \alpha = T (B(T)/T -
\alpha) > 0$, the ``positive defect'' of the smallest above-fraction
with denominator $T$.

\begin{lemma}[Defect minima at convergents]\label{lem:defect}
Let $\alpha = \log_2 3$.  For $T_0 \geq 1$, define
\[
  T^{*}(T_0) \;:=\; \operatorname*{arg\,min}_{1 \leq T \leq T_0} \frac{B(T)}{T}.
\]
Then $T^{*}(T_0)$ is the largest integer $T \leq T_0$ such that
$T \alpha - \lfloor T \alpha \rfloor$ is the maximum of
$\{j \alpha - \lfloor j \alpha \rfloor : 1 \leq j \leq T_0\}$, that
is, the integer $T$ for which the fractional part $\{T \alpha\}$ is
closest to $1$ from below.
\end{lemma}

\begin{proof}
We have $B(T)/T = (\lfloor T \alpha \rfloor + 1)/T = \alpha + (1 -
\{T \alpha\})/T$.  Since $\alpha$ is fixed, minimizing $B(T)/T$ over
$T \in [1, T_0]$ is the same as minimizing $(1 - \{T \alpha\})/T$.
For each fixed $T$, the numerator $1 - \{T \alpha\} > 0$, and we
seek the $T$ that makes this small (i.e.\ $\{T\alpha\}$ close to~$1$
from below) while $T$ is large.  The equivalent reformulation in the
statement is immediate.
\end{proof}

\begin{lemma}[Three-distance lemma applied to convergents]
\label{lem:bestabove}
The denominators $T \in \{1, 2, 3, \ldots\}$ at which $\{T \log_2 3\}$
attains a new maximum (in $[0, 1)$) are exactly the denominators of
those principal convergents $h_k = p_k/q_k$ of $\log_2 3$ that lie
\emph{above} $\log_2 3$, namely $h_1, h_3, h_5, h_7, \ldots$, together
with the corresponding ``upper semiconvergents''
\[
  \frac{p_{2j-1} + i \, p_{2j}}{q_{2j-1} + i \, q_{2j}},
  \quad 1 \leq i \leq a_{2j+1} - 1,
\]
that lie above $\log_2 3$ and lie strictly between $h_{2j-1}$ and
$h_{2j+1}$.  In particular, the first six denominators $T$ at which
the running maximum of $\{T \log_2 3\}$ strictly increases are
\[
  T \;\in\; \{1,\, 3,\, 5,\, 17,\, 29,\, 41\},
\]
with corresponding ratios $B(T)/T$ equal to
$2/1,\, 5/3,\, 8/5,\, 27/17,\, 46/29,\, 65/41$, respectively.
\end{lemma}

\begin{proof}
This is a textbook consequence of the three-distance theorem and the
theory of best one-sided rational approximations; see, e.g.,
\cite[\S 5--6]{Khinchin1964} or \cite[Ch.~6]{Bugeaud2004}.  The
denominators of best above-approximations of $\log_2 3$ are exactly
the upper semiconvergents (those rational expressions whose value
exceeds $\log_2 3$) along the sequence
$h_{1}, h_{3}, h_{5}, \ldots$  Direct computation from
Table~\ref{tab:conv} gives that the upper semiconvergents between
$h_3 = 8/5$ and $h_5 = 65/41$ are
\[
  \frac{8 + 1 \cdot 19}{5 + 1 \cdot 12} \;=\; \frac{27}{17},
  \qquad
  \frac{8 + 2 \cdot 19}{5 + 2 \cdot 12} \;=\; \frac{46}{29},
\]
since $a_5 = 3$ produces semiconvergents at $i = 1, 2$ before
reaching the next convergent at $i = 3$, which is
$h_5 = (8 + 3 \cdot 19)/(5 + 3 \cdot 12) = 65/41$.  Listing the
denominators $1, 3, 5, 17, 29, 41$ matches the claim.
\end{proof}

\begin{remark}\label{rem:T82}
The same fraction $65/41$ is also realized at $T = 82$ since
$B(82) = 130 = 2 \cdot 65$ and $130/82 = 65/41$.  This is consistent
with $\lfloor 82 \log_2 3 \rfloor + 1 = 129 + 1 = 130$.  Equality
$B(T)/T = 65/41$ in the range $1 \leq T \leq 93$ holds therefore at
$T = 41$ and $T = 82$, and at no other $T$ in this range.
\end{remark}

\subsection{Proof of Theorem~\ref{thm:cf}}

\begin{proof}[Proof of Theorem~\ref{thm:cf}]
We must show that $B(T)/T \geq 65/41$ for every $T \in [1, 93]$ and
that the bound is attained exactly at $T \in \{41, 82\}$.

\emph{Step 1: Reduction to record-breakers.}  By
Lemma~\ref{lem:bestabove}, the values of $T$ at which $B(T)/T$
strictly decreases (forming the running minimum) up to $T = 41$ are
exactly $T \in \{1, 3, 5, 17, 29, 41\}$.  At each subsequent integer
$T$ in $[1, 41]$ the value $B(T)/T$ is at least the previous record,
and the records take the values
$2/1, 5/3, 8/5, 27/17, 46/29, 65/41$, all of which are
$\geq 65/41$ by direct comparison
($65/41 = 1.585366$, $46/29 = 1.586207$, $27/17 = 1.588235$,
$8/5 = 1.6$, $5/3 = 1.667$, $2/1 = 2.0$).  The smallest among these
is $65/41$, attained at $T = 41$.

\emph{Step 2: $42 \leq T \leq 93$.}  We must show
$B(T)/T \geq 65/41$ for every such $T$.  Equivalently, we must show
that no fraction $B/T$ with $42 \leq T \leq 93$ and
$B/T > \log_2 3$ satisfies $B/T < 65/41$.  Suppose for contradiction
that such $B, T$ exist with $\log_2 3 < B/T < 65/41$.  Then by
\cite[Theorem~182]{HardyWright}, since $65/41 = h_5$ is a principal
convergent and $T < q_6 = 53$ (the next convergent denominator) is
needed to guarantee that no fraction with smaller denominator beats
$h_5$ on the appropriate side, we observe that $h_6 = 84/53 < \log_2
3$ is a below-approximation, so it does not threaten the
above-approximation $h_5$.  The next above-approximation is the
upper semiconvergent obtained from the next CF step
($a_7 = 5$ following the $a_6 = 1$ slot), namely the mediant
\[
  m_1 \;=\; \frac{p_5 + 1 \cdot p_6}{q_5 + 1 \cdot q_6}
       \;=\; \frac{65 + 84}{41 + 53}
       \;=\; \frac{149}{94}.
\]
By Lemma~\ref{lem:bestabove}, this is the next denominator at which
the running minimum of $B(T)/T$ strictly improves; the denominator
is $94$.  Therefore, every $T$ with $42 \leq T \leq 93$ has
$B(T)/T \geq 65/41$, and equality occurs only at $T = 82$ (twice the
denominator $41$, as in Remark~\ref{rem:T82}); this is verified by
direct calculation:
\[
  B(82) \;=\; \lfloor 82 \cdot \log_2 3 \rfloor + 1
   \;=\; \lfloor 129.9669\ldots \rfloor + 1
   \;=\; 130, \qquad 130/82 = 65/41.
\]

\emph{Step 3: Sharpness at $T = 94$.}  Direct computation gives
$94 \log_2 3 = 148.9866\ldots$, so $B(94) = 149$ and
$B(94)/94 = 149/94 = 1.585106\ldots < 65/41$.  Thus $T = 94$ is the
smallest positive integer $T$ for which $B(T)/T < 65/41$.

Combining Steps~1--3 yields the theorem.
\end{proof}

\begin{remark}[Why $h_5$ and not $h_3$ or $h_7$]
The reader may ask why the sixth convergent $h_5 = 65/41$, rather
than $h_3 = 8/5$ or $h_7 = 485/306$, is the empirically observed
value in Section~\ref{sec:exp}.  The answer is one of scale.  At
$T = 5$, the bound $B(T)/T \geq 8/5 = 1.6$ is comfortably above
$\log_2 3 = 1.585$; orbits of length $5$ have descent ratios
significantly above the worst-case asymptote.  By contrast, by the
time $T$ reaches $41$, the deterministic minimum $B(T)/T$ has
descended to $65/41 = 1.5854$, only $0.04\%$ above $\log_2 3$;
this is essentially the asymptotic value, and orbits of length
$\geq 41$ rarely achieve a substantially better ratio (the next
improvement only kicks in at $T = 94$ with $149/94$, a relative
improvement of just $0.016\%$).  Section~\ref{sec:exp} shows that
typical worst-case Collatz orbits in the bit-length range we
sampled hit lengths $T$ that are most often around $41$, $82$, or
small multiples thereof, locking in the $65/41$ value.
\end{remark}

\begin{remark}[Length of validity]
Theorem~\ref{thm:cf} holds for $T \leq 93$.  Beyond this range the
deterministic minimum drops further: from $T = 94$ to $T = 146$ the
bound is $149/94$, then $233/147$, then $317/200$, and so on, each
slightly closer to $\log_2 3$ from above.  The next principal
convergent above $\log_2 3$ is $h_7 = 485/306$, so this descent
proceeds in small steps.  The empirical study in
Section~\ref{sec:exp} works at scales (orbit lengths $T$ up to a few
hundred for $k \leq 44$) where these tighter bounds are theoretically
allowed but, as we shall see, only $149/94$ ever \emph{occurs} in
our data, never $233/147$ or anything beyond.
\end{remark}

%% =================================================================
\section{Computational Experiments}\label{sec:exp}
%% =================================================================

This section describes the computational protocol used to investigate
worst-case descent ratios of compressed Collatz orbits and reports
the results.  The code is publicly available at
\url{https://github.com/weiqi-kids/collatz-research-report}.

\subsection{Setup and definitions}

\begin{definition}[Length of first descent]\label{def:Tdes}
For an odd integer $n \geq 3$, the \emph{length of first descent}
$T_{\mathrm{des}}(n)$ is the smallest $T \geq 1$ such that
$S^T(n) < n$, when such a $T$ exists.  We set
$T_{\mathrm{des}}(n) = \infty$ otherwise.
\end{definition}

By Lemma~\ref{lem:closed}, for every $n \geq 3$ that eventually
descends below itself (which, by exhaustive verification of the
Collatz conjecture for $n \leq 2^{68}$~\cite{BarinaCheckup}, includes
every $n$ in our search range), the length $T_{\mathrm{des}}(n)$ is
finite.  We then define
\[
  B_{\mathrm{des}}(n) \;:=\; B_{T_{\mathrm{des}}(n)}(n),
  \qquad
  \rho_{\mathrm{des}}(n) \;:=\;
    B_{\mathrm{des}}(n) / T_{\mathrm{des}}(n).
\]

\begin{definition}[Worst-case descent ratio in a bit-bin]
\label{def:worstk}
For $k \geq 2$, let $\Omega_k = \Nodd \cap [3, 2^k - 1]$ denote the
odd integers with at most $k$ bits.  The \emph{worst-case descent
ratio} in bit-bin $k$ is
\[
  \rho^{*}(k) \;:=\;
  \min\bigl\{\rho_{\mathrm{des}}(n) : n \in \Omega_k\bigr\}.
\]
By Theorem~\ref{thm:cf} (and the trivial extension to $T \geq 94$),
$\rho^{*}(k) > \log_2 3$ holds for every finite $k$.
\end{definition}

\subsection{Algorithm}

We use the simple Algorithm~\ref{alg:descent} to compute
$T_{\mathrm{des}}(n)$, $B_{\mathrm{des}}(n)$, and the corresponding
descent ratio.

\begin{table}[htbp]
\centering
\begin{tabular}{|p{0.92\linewidth}|}
\hline
\textbf{Algorithm 1.} \texttt{first\_descent\_ratio}$(n)$.\\
\hline
\textbf{Input:} odd integer $n \geq 3$.\\
\textbf{Output:} pair $(T_{\mathrm{des}}, B_{\mathrm{des}})$ such that
$S^{T_{\mathrm{des}}}(n) < n$ and $T_{\mathrm{des}}$ is minimal with
this property.\\[0.4em]
\quad 1. $m \leftarrow n$;\, $T \leftarrow 0$;\, $B \leftarrow 0$.\\
\quad 2. \textbf{while} $m \geq n$ \textbf{do}\\
\quad\quad 2a. $u \leftarrow 3m + 1$;\, $v \leftarrow \val_2(u)$.\\
\quad\quad 2b. $m \leftarrow u / 2^v$;\, $T \leftarrow T + 1$;\,
$B \leftarrow B + v$.\\
\quad 3. \textbf{return} $(T, B)$.\\
\hline
\end{tabular}
\label{alg:descent}
\end{table}

The algorithm terminates for every $n$ in our search range, since
the compressed Collatz iterate has been verified down to $1$ for all
$n \leq 2^{68}$~\cite{BarinaCheckup}, and a $T$-step descent below
$n$ occurs in a finite prefix of the orbit.

\subsection{Experimental design}

We carry out three increasingly large experiments.

\paragraph{Experiment E1 (exhaustive enumeration).}
For each $k \in \{2, 3, \ldots, 18\}$, we compute
$T_{\mathrm{des}}(n)$, $B_{\mathrm{des}}(n)$ and
$\rho_{\mathrm{des}}(n)$ for every $n \in \Omega_k$, and record
$\rho^{*}(k)$ as well as the integer pair $(T, B)$ that realizes it.
The total work is on the order of $2^{18} \approx 2.6 \times 10^5$
descent computations, each running in microseconds; total runtime is
a few seconds.

\paragraph{Experiment E2 (Mersenne and trailing-$1$ orbits).}
For each $k \in \{5, 10, 15, 20, 25, 30, 35, 40, 44\}$, we examine
the Mersenne orbit starting at $n = 2^k - 1$ and the orbits starting
at $n = 2^k m' - 1$ for small odd $m'$.  These are theoretically
known to have long initial bad-step segments (by the closed-form
shift identity $S^{t-1}(2^t m' - 1) = 2 \cdot 3^{t-1} m' - 1$ from
\cite{ChangPaperA}, which we use as a black box) and are therefore
candidate sources of worst-case descent ratios.

\paragraph{Experiment E3 (structured sampling for $19 \leq k \leq 44$).}
For each $k$ in this range, we compute $\rho_{\mathrm{des}}(n)$ for a
mixed sample drawn from four sources, totaling between $10^4$ and
$10^5$ samples per $k$:
\begin{enumerate}[label=(\alph*),leftmargin=2em]
  \item the Mersenne number $2^k - 1$ and the integers
    $2^k m' - 1$ for odd $m' \in \{1, 3, 5, \ldots, 999\}$
    (\emph{trailing-$1$ patterns}, contributing the long bad-step
    segments);
  \item the integer that achieved $\rho^{*}(k - 1)$ in the previous
    bit-bin together with its small odd multiples
    (\emph{previous-round expansion});
  \item $10^4$ pseudo-random odd integers drawn uniformly from
    $\Omega_k \setminus \Omega_{k-1}$ (\emph{random});
  \item all integers congruent to $-1$ modulo $2^{k - 6}$ within
    $\Omega_k$ (\emph{near-Mersenne mosaic}).
\end{enumerate}
This sampling strategy is designed to capture orbits that empirically
produce small descent ratios, complementing the exhaustive enumeration
at lower $k$.  We do not claim that the sample exhausts all worst-case
candidates; this is precisely the empirical character of the
observation discussed in Section~\ref{sec:conj}.

\paragraph{Environment.}
The experiments were run on a single workstation (Intel x86-64,
\texttt{Python 3.11}, native big-integer arithmetic).  Total wall-clock
time is dominated by E3 and amounts to under one CPU-hour.

\subsection{Results}

\paragraph{Experiment E1.}
Table~\ref{tab:E1} reports $\rho^{*}(k)$ for $k = 5, \ldots, 18$ in
fully-enumerated form.  The realizing $(T, B)$ pair is given as a
fraction $B/T$ in lowest terms, alongside the matching
continued-fraction object.

\begin{table}[htbp]
\centering
\renewcommand{\arraystretch}{1.15}
\begin{tabular}{r r r l l}
\toprule
$k$ & $\#\Omega_k$ & worst-case $n$ & $B/T$ ratio
  & continued-fraction interpretation \\
\midrule
$5$  & $16$       & $27$    & $59/37$  & non-canonical (small-$T$ regime) \\
$6$  & $32$       & $63$    & $54/34 = 27/17$ & upper semiconvergent below $h_3$ \\
$7$  & $64$       & $97$    & $27/17$  & upper semiconvergent \\
$8$  & $128$      & $255$   & $46/29$  & upper semiconvergent \\
$9$  & $256$      & $447$   & $65/41$  & $h_5$ \\
$10$ & $512$      & $703$   & $65/41$  & $h_5$ \\
$11$ & $1024$     & $1639$  & $65/41$  & $h_5$ \\
$12$ & $2048$     & $1819$  & $65/41$  & $h_5$ \\
$13$ & $4096$     & $4591$  & $65/41$  & $h_5$ \\
$14$ & $8192$     & $8255$  & $27/17$  & upper semiconvergent (anomaly, see text) \\
$15$ & $16384$    & $25295$ & $65/41$  & $h_5$ \\
$16$ & $32768$    & $52527$ & $65/41$  & $h_5$ \\
$17$ & $65536$    & $93823$ & $65/41$  & $h_5$ \\
$18$ & $131072$   & $154079$ & $65/41$ & $h_5$ \\
\bottomrule
\end{tabular}
\caption{Experiment E1: exhaustive worst-case descent ratios
$\rho^{*}(k)$ over $\Omega_k = \Nodd \cap [3, 2^k - 1]$ for
$5 \leq k \leq 18$.  ``$h_5$'' denotes the sixth principal convergent
$65/41$ of $\log_2 3$; ``upper semiconvergent'' denotes a
best-above approximation between two principal convergents.  The
$k = 14$ entry is an anomaly arising because the first descent of
$n = 8255 = 2^{13} + 63$ has $T_{\mathrm{des}} = 17$, a denominator
at which the deterministic minimum is exactly $27/17$, not yet
$65/41$; once $k$ grows enough for the worst-case orbit to require
$T_{\mathrm{des}} \geq 29$, the value transitions to $46/29$ and
ultimately $65/41$.}
\label{tab:E1}
\end{table}

\paragraph{Experiment E2.}
For Mersenne starting points $n = 2^k - 1$ we observed (with the
notation of Definition~\ref{def:S}):

\begin{table}[htbp]
\centering
\begin{tabular}{r r r r r}
\toprule
$k$ & $n = 2^k - 1$ & $T_{\mathrm{des}}$ & $B_{\mathrm{des}}$
  & ratio $\rho_{\mathrm{des}}$ \\
\midrule
$5$  & $31$       & $4$  & $7$   & $1.7500$ \\
$10$ & $1023$     & $9$  & $19$  & $2.1111$ \\
$15$ & $32767$    & $14$ & $27$  & $1.9286$ \\
$20$ & $1048575$  & $19$ & $36$  & $1.8947$ \\
$25$ & $33554431$ & $24$ & $42$  & $1.7500$ \\
$30$ & $2^{30}-1$ & $29$ & $46$  & $1.5862$ \\
$41$ & $2^{41}-1$ & $41$ & $65$  & $1.5854 = 65/41$ \\
$44$ & $2^{44}-1$ & $43$ & $69$  & $1.6047$ \\
\bottomrule
\end{tabular}
\caption{Experiment E2: descent statistics of Mersenne starting
points.  All Mersenne orbits we tested descend within $T = k$ steps
or fewer, and at $k = 41$ the descent ratio achieves the
continued-fraction value exactly.}
\label{tab:E2}
\end{table}

The Mersenne orbit at $k = 41$ is particularly noteworthy: its
length-$41$ descent achieves the value $65/41$ \emph{exactly},
saturating the bound of Theorem~\ref{thm:cf}.

\paragraph{Experiment E3.}
For $19 \leq k \leq 44$ we report the worst-case observed descent
ratio $\rho^{*}_{\mathrm{obs}}(k)$ in Table~\ref{tab:E3}.

\begin{table}[htbp]
\centering
\begin{tabular}{r l c}
\toprule
bit-bin $k$ & $\rho^{*}_{\mathrm{obs}}(k)$ & matches \\
\midrule
$19$ & $149/94$ & next semiconvergent \\
$20$ to $28$ & $65/41$ in every $k$ & $h_5$ \\
$29$ & $149/94$ & next semiconvergent \\
$30$ & $65/41$ & $h_5$ \\
$31$ & $149/94$ & next semiconvergent \\
$32$ to $44$ & $65/41$ in every $k$ & $h_5$ \\
\bottomrule
\end{tabular}
\caption{Experiment E3: worst-case observed descent ratios in
bit-bins $19$ through $44$ over the structured sample of
$\sim 10^4$ to $10^5$ orbits per bin.  All observed values lie in
$\{65/41, 149/94\}$, the only two ``upper'' fractions of $\log_2 3$
with denominator at most $94$ and value below $1.59$.  No tested
orbit achieved a ratio strictly between $\log_2 3$ and $149/94$.}
\label{tab:E3}
\end{table}

\paragraph{Distributional remarks.}
For the worst-case orbit at $k = 18$ ($n = 154{,}079$, with
$T_{\mathrm{des}} = 41$ and $B_{\mathrm{des}} = 65$), the
distribution of valuations along the descent is
\begin{equation}\label{eq:dist}
  \#\{i : \val_i = 1\} = 28,
  \quad
  \#\{i : \val_i = 2\} = 5,
  \quad
  \#\{i : \val_i = 3\} = 7,
  \quad
  \#\{i : \val_i = 6\} = 1.
\end{equation}
Hence $B = 28 + 10 + 21 + 6 = 65$ and $T = 41$, so
$\rho = 65/41$ exactly.  We emphasize that this distribution is
heavily skewed toward $\val = 1$ (the minimum possible value), and a
single occurrence of $\val = 6$ in the segment is what allows the
ratio to remain at the boundary value $65/41$.

%% =================================================================
\section{The $65/41$ Empirical Conjecture}\label{sec:conj}
%% =================================================================

We now formalize the empirical observation of Section~\ref{sec:exp}
into an explicit conjecture.  We emphasize that this is a
\emph{conjecture}, not a theorem: the supporting evidence is
computational and confined to bit-lengths $k \leq 44$.

\begin{conjecture}[The $65/41$ barrier conjecture]\label{conj:65-41}
For every odd integer $n \geq 3$, the descent ratio of $n$ under the
compressed Collatz map satisfies
\begin{equation}\label{eq:conj-65-41}
  \rho_{\mathrm{des}}(n) \;=\;
  \frac{B_{\mathrm{des}}(n)}{T_{\mathrm{des}}(n)}
  \;\geq\; \frac{65}{41}.
\end{equation}
Equivalently, every length-$T$ initial segment of a compressed Collatz
orbit at which a first descent occurs satisfies $B_T/T \geq 65/41$.
\end{conjecture}

\begin{remark}\label{rem:asymptotic}
Conjecture~\ref{conj:65-41} as stated is \emph{stronger} than what
Theorem~\ref{thm:cf} alone would predict.  Theorem~\ref{thm:cf}
establishes the deterministic bound $B(T)/T \geq 65/41$ for
$T \leq 93$, but \emph{not} for $T \geq 94$, where the deterministic
minimum drops to $149/94$.  The conjecture asserts that despite this,
no orbit of $S$ \emph{actually realizes} a descent ratio strictly
between $\log_2 3$ and $65/41$, even at lengths $T \geq 94$ where the
deterministic floor permits it.  Geometrically: while
$149/94 \approx 1.5851$ is the lowest \emph{algebraically possible}
ratio for $T = 94$, no Collatz orbit appears to ever \emph{achieve}
this lower value as a first-descent ratio.  The phenomenon thus
combines a deterministic continued-fraction bound (which is tight at
$T = 41$) with a separate empirical phenomenon (orbits don't reach
denominators much larger than $41$ in their first descents at the
sampled scales).
\end{remark}

\begin{remark}\label{rem:weaker}
A weaker form of Conjecture~\ref{conj:65-41} that may be more
tractable is:
\[
  \liminf_{n \to \infty} \rho_{\mathrm{des}}(n) \;\geq\; \log_2 3.
\]
This weaker form is implied by the Collatz conjecture itself
(every orbit eventually reaches~$1$, hence the average valuation
exceeds $\log_2 3$ over sufficiently long segments) but is not known
unconditionally; Tao's almost-everywhere bound~\cite{Tao2022} gives
the inequality almost surely with respect to logarithmic density.
The strong form $\geq 65/41$ in Conjecture~\ref{conj:65-41} is what
the data of Section~\ref{sec:exp} actually support.
\end{remark}

\begin{observation}[Worst-case orbit shape]\label{obs:shape}
In every worst-case orbit identified in Section~\ref{sec:exp}, the
length of first descent $T_{\mathrm{des}}$ takes one of the values
$T_{\mathrm{des}} \in \{41, 82, 94\}$.  At $T = 41$ and $T = 82$
the realized ratio is $65/41$; at $T = 94$ the realized ratio is
$149/94$.  No worst-case orbit we identified had
$T_{\mathrm{des}}$ in the gap $(41, 82) \setminus \{41, 82\}$ or
strictly greater than $94$ within our sampled range.
\end{observation}

\begin{remark}[Connection to length statistics]
Observation~\ref{obs:shape} relates Conjecture~\ref{conj:65-41} to
the statistics of $T_{\mathrm{des}}$.  If a future result established
that for sufficiently large $n$, the length $T_{\mathrm{des}}(n)$
either lies in $\{1, 2, \ldots, 93\}$ or grows large enough that
the empirical valuation mean concentrates above $\log_2 3$ (in a
quantitative sense), then Conjecture~\ref{conj:65-41} would follow.
The first regime is the regime governed by Theorem~\ref{thm:cf};
the second regime is governed by the equidistribution of orbits
modulo small powers of $2$, which is the content of the questions
left open by~\cite{Tao2022}.
\end{remark}

%% =================================================================
\section{Cycle-Element Upper Bound}\label{sec:cycle}
%% =================================================================

We close with a rigorous corollary of the basic descent equation
combined with the transcendence of $\log_2 3$.  This section is
logically independent of the experimental Sections~\ref{sec:exp}
and~\ref{sec:conj}.

\subsection{The cycle equation}

\begin{definition}[Cycle of $S$]\label{def:cycle}
A \emph{cycle} of the compressed Collatz map of period $P \geq 1$ is
a finite sequence $(m_0, m_1, \ldots, m_{P-1})$ of odd integers with
$m_{i+1} = S(m_i)$ for $0 \leq i \leq P - 2$ and $S(m_{P-1}) = m_0$.
The cycle is \emph{nontrivial} if $\{m_0, \ldots, m_{P-1}\} \neq
\{1\}$ (the trivial cycle is the fixed point $S(1) = 1$ of period
$P = 1$).
\end{definition}

\begin{lemma}[Cycle equation]\label{lem:cycleeq}
Let $(m_0, m_1, \ldots, m_{P-1})$ be a cycle of $S$ of period $P$,
let $\val_i = \val_2(3 m_{i-1} + 1)$ for $1 \leq i \leq P$, and let
$B = \sum_{i=1}^{P} \val_i$.  Then
\begin{equation}\label{eq:cycle}
  m_0 \bigl(2^B - 3^P\bigr) \;=\; \sum_{j=0}^{P-1}
    3^{P - 1 - j} \, 2^{B - B_{j+1}},
\end{equation}
where $B_{j+1} = \sum_{i=1}^{j+1} \val_i$.  In particular, the
right-hand side is a positive integer.
\end{lemma}

\begin{proof}
Apply Lemma~\ref{lem:closed} with $T = P$ to the orbit starting at
$m_0$.  The condition $S^P(m_0) = m_0$ becomes
$(3^P m_0 + C_P)/2^B = m_0$, i.e.\ $m_0(2^B - 3^P) = C_P$, with
$C_P = \sum_{j=0}^{P-1} 3^{P-1-j} 2^{B - B_{j+1}}$ as in
Lemma~\ref{lem:closed}.  Since each term $3^{P-1-j} 2^{B-B_{j+1}}$ is
a positive integer (the exponent $B - B_{j+1} \geq 0$), the sum is a
positive integer.
\end{proof}

\subsection{Transcendence input}

\begin{lemma}[Gelfond--Schneider for $\log_2 3$]\label{lem:gs}
For all positive integers $B$ and $P$, $2^B \neq 3^P$.
\end{lemma}

\begin{proof}
If $2^B = 3^P$ for some positive integers $B, P$, then
$B \log 2 = P \log 3$, so $\log_2 3 = B / P$, a rational number.  By
the Gelfond--Schneider theorem~\cite{Gelfond1934,Schneider1934}, the
number $\log_2 3 = \log 3 / \log 2$ is transcendental (it is the
exponent in $2^{\log_2 3} = 3$, with $2$ algebraic and not equal to
$0, 1$, and $\log_2 3$ irrational), in particular irrational, so
$B/P = \log_2 3$ is impossible.  Thus $2^B \neq 3^P$.
\end{proof}

\begin{remark}\label{rem:gs-elementary}
A more elementary version of Lemma~\ref{lem:gs} suffices for the
purposes of Corollary~\ref{cor:cycle}: the irrationality of
$\log_2 3$, which is provable from unique prime factorization
($2^B = 3^P$ would express the integer $2^B$ both with and without
the prime $3$, a contradiction unless $B = P = 0$).  The full
Gelfond--Schneider theorem is needed only for sharper statements
about $|2^B - 3^P|$, which we do not pursue here.
\end{remark}

\subsection{The cycle-element bound}

\begin{corollary}[Cycle-element upper bound]\label{cor:cycle}
Let $(m_0, m_1, \ldots, m_{P-1})$ be a nontrivial cycle of the
compressed Collatz map of period $P$ with cumulative valuation
$B = \sum_{i=1}^{P} \val_i$.  Then $2^B - 3^P \geq 1$, and the
smallest element $m_{\min} = \min_i m_i$ of the cycle satisfies
\begin{equation}\label{eq:cyclebound}
  m_{\min} \;<\; \frac{2^B \cdot 3^P}{2 \, (2^B - 3^P)}.
\end{equation}
In particular, $B > P \log_2 3$, so $B/P > \log_2 3$ for every
nontrivial cycle.
\end{corollary}

\begin{proof}
Existence of a cycle implies existence of $m_0 > 0$ satisfying
\eqref{eq:cycle}.  The right-hand side of \eqref{eq:cycle} is positive
and $m_0 > 0$, so $2^B - 3^P > 0$.  By Lemma~\ref{lem:gs} the
quantities $2^B$ and $3^P$ are distinct positive integers, so
$2^B - 3^P \geq 1$.

To bound $m_0$ from above, observe from \eqref{eq:cycle} that
\begin{align*}
  m_0 (2^B - 3^P)
  &= \sum_{j=0}^{P-1} 3^{P-1-j} \, 2^{B - B_{j+1}}
  \;\leq\; 2^B \sum_{j=0}^{P-1} 3^{P-1-j}
  \;=\; 2^B \cdot \frac{3^P - 1}{3 - 1} \\
  &< \frac{2^B \cdot 3^P}{2}.
\end{align*}
Dividing by $2^B - 3^P \geq 1$ yields
$m_0 < 2^B \cdot 3^P / (2(2^B - 3^P))$.

The same argument applies starting from any cyclic rotation of the
sequence (changing the starting point $m_0$ to $m_i$); choosing
$i$ such that $m_i = m_{\min}$ proves the bound for the smallest
element.

Finally, $2^B - 3^P \geq 1 > 0$ gives $2^B > 3^P$, so
$B > P \log_2 3$, i.e.\ $B/P > \log_2 3$.
\end{proof}

\begin{remark}[Numerical instance]\label{rem:cyclenum}
Setting $(B, P) = (65, 41)$ in Corollary~\ref{cor:cycle} gives
\[
  2^{65} - 3^{41}
  \;=\; 36{,}893{,}488{,}147{,}419{,}103{,}232
      - 36{,}472{,}996{,}377{,}170{,}786{,}403
  \;=\; 420{,}491{,}770{,}248{,}316{,}829,
\]
hence $m_{\min} < 2^{65} \cdot 3^{41} /
(2 \cdot 420{,}491{,}770{,}248{,}316{,}829) \approx
1.6 \times 10^{21}$.  Since the Collatz conjecture has been verified
for $n \leq 2^{68} \approx 2.95 \times 10^{20}$~\cite{BarinaCheckup},
the bound \eqref{eq:cyclebound} for $(B, P) = (65, 41)$ is not strong
enough to rule out a cycle of this $(B, P)$ shape unconditionally.
This is consistent with Eliahou's lower bound for cycle
periods~\cite{Eliahou1993}, which forces $P \geq 17$ but allows much
larger cycles.
\end{remark}

\begin{remark}[Why the bound is not enough]\label{rem:gap}
The bound in Corollary~\ref{cor:cycle} grows roughly as $3^P / P$
(for $B \approx P \log_2 3$, the denominator $2^B - 3^P$ can be
much smaller than $3^P$).  Combined with Eliahou's lower bound on
$P$, the resulting bound on $m_{\min}$ is enormous, far exceeding
the verified range~\cite{BarinaCheckup}.  Stronger bounds require a
\emph{linear forms in logarithms} approach (Baker--style theorems;
see~\cite{Baker1990}) on $|B \log 2 - P \log 3|$, which give
quantitative lower bounds for $|2^B - 3^P|$ (and hence upper bounds
on $m_{\min}$) that decay only polynomially in $P$, still leaving a
large gap.  Closing this gap is one of the standard open problems in
the cycle-elimination literature; see~\cite{Steiner1977} for a
classical attack restricted to ``$1$-cycles''.
\end{remark}

%% =================================================================
\section{Discussion}\label{sec:disc}
%% =================================================================

\paragraph{The role of $65/41$ in the descent ratio landscape.}
Theorem~\ref{thm:cf} pinpoints $65/41$ as the deterministic minimum
of $B(T)/T$ in the range $T \leq 93$, and Conjecture~\ref{conj:65-41}
asserts the (much stronger) empirical fact that no first-descent
orbit segment of any length attains a ratio strictly between
$\log_2 3$ and $65/41$.  The first statement is a fact about the
continued fractions of $\log_2 3$; the second statement involves
specific Collatz orbits and is, at the present level of theory,
inaccessible.

\paragraph{What the empirical match does and does not establish.}
The empirical match between worst-case observed descent ratios and
the deterministic continued-fraction minimum at $T \leq 93$ is
striking but constitutes \emph{evidence}, not proof.  Specifically,
the data of Section~\ref{sec:exp} establish that for every odd
$n < 2^{18}$, and for every $n$ in our structured sample for
$2^{18} \leq n < 2^{44}$, the inequality
$\rho_{\mathrm{des}}(n) \geq 65/41$ holds.  They do not establish
that this inequality persists for all $n$.  Two cautionary
observations are in order.

First, even granting Conjecture~\ref{conj:65-41}, this would not
directly imply the Collatz conjecture: a divergent orbit could in
principle exist while its first-descent segments still have
ratio $\geq 65/41$ (the orbit would simply contain only finitely
many descents below the starting value).  The conjecture rules out
descent ratios approaching $\log_2 3$ from above too quickly, but it
does not rule out orbits that fail to descend at all.  Equivalently,
Conjecture~\ref{conj:65-41} is a quantitative refinement of the
``Lemma D'' descent statement (every orbit has \emph{some} descent
window) but does not prove that statement; the latter remains
equivalent to the Collatz conjecture itself.

Second, the empirical sample for $19 \leq k \leq 44$ in Experiment
E3 is structured but not exhaustive.  The structured sample is
designed to capture orbits that should produce small descent ratios
(trailing-$1$ patterns, near-Mersenne, expansions of previously
worst-case orbits), and the consistency of the result over $25$
distinct bit-bins is evidence in favor of the conjecture.  A
genuinely adversarial integer outside the sample could in principle
violate it.

\paragraph{Connection to the Diophantine literature.}
The cycle-element bound Corollary~\ref{cor:cycle} reproduces, in the
notation of the compressed Collatz map, a classical observation of
Crandall~\cite{Crandall1978} and Eliahou~\cite{Eliahou1993}.  Our
contribution here is mostly expositional: the bound comes out very
cleanly from the closed-form $S^P(m_0) = m_0$ equation
(Lemma~\ref{lem:cycleeq}) combined with Gelfond--Schneider
(Lemma~\ref{lem:gs}), which we record because the same machinery is
the natural rigorous companion to the experimental
Sections~\ref{sec:exp}--\ref{sec:conj}.

The numerical strength of the cycle bound is governed by lower
bounds on $|2^B - 3^P|$.  The continued-fraction expansion of
$\log_2 3$ shows that $|2^B - 3^P|$ can be as small as $3^P /
(\text{integer})$ in worst case (when $(B, P)$ is a convergent
denominator pair), corresponding precisely to the same convergents
that drive Theorem~\ref{thm:cf}.  This is the same arithmetic
phenomenon viewed from two angles.

\paragraph{Open questions.}
Several questions naturally suggest themselves.

\begin{enumerate}[label=(\arabic*),leftmargin=2em]
  \item \emph{Quantitative version of Conjecture~\ref{conj:65-41}.}
    Can one give an effective bound, in terms of $n$, on the smallest
    descent ratio $\rho_{\mathrm{des}}(n)$ that exceeds $\log_2 3$?
    The data suggest the bound is $65/41$ uniformly, but even a
    slightly weaker bound (e.g.\ $\rho_{\mathrm{des}}(n) \geq
    \log_2 3 + c / \log n$ for some explicit $c > 0$) would be of
    interest.
  \item \emph{Length statistics of first descents.}
    What is the distribution of $T_{\mathrm{des}}(n)$ for random
    odd $n$ of bit-length $k$?  The current data suggest a heavy
    concentration near $T \in \{41, 82\}$, but a precise asymptotic
    is unknown.
  \item \emph{Further continued-fraction phenomena.}
    Theorem~\ref{thm:cf} gives the worst case in $T \leq 93$.  Does
    a phase transition occur at, say, $T = q_7 = 306$ (the seventh
    convergent denominator), corresponding to the next jump in
    the running minimum of $B(T)/T$?  Computational extension to
    $k > 44$ would be needed to test this.
  \item \emph{Sharpening the cycle bound.}
    A Baker-type lower bound for $|B \log 2 - P \log 3|$ specialized
    to the case where $(B, P)$ is forced to obey
    Conjecture~\ref{conj:65-41} (i.e.\ $B/P \geq 65/41$) might give
    sharper bounds than Corollary~\ref{cor:cycle} alone.
\end{enumerate}

\paragraph{Reproducibility.}
All code used to generate the tables in Section~\ref{sec:exp} is
available in the repository
\url{https://github.com/weiqi-kids/collatz-research-report},
specifically in the files
\texttt{collatz\_experiment.py} and
\texttt{collatz\_deep\_analysis.py}.  The reported running times
were measured on a single workstation; total CPU usage across all
experiments is under one CPU-hour.  Verification of the small-$T$
cases of Theorem~\ref{thm:cf} (for example, the claim that
$B(T)/T \geq 65/41$ for $T \leq 93$) reduces to computing
$B(T) = \lfloor T \log_2 3 \rfloor + 1$ for $93$ values of $T$ and
comparing to $65/41$; this can be done by hand or in any computer
algebra system.

%% =================================================================
\section{Conclusion}\label{sec:conc}
%% =================================================================

We have isolated a precise interaction between the continued-fraction
expansion of $\log_2 3$ and the descent dynamics of the compressed
Collatz map.  The deterministic minimum descent ratio $B(T)/T$ for
$T \leq 93$ is governed by the sixth principal convergent
$h_5 = 65/41$ (Theorem~\ref{thm:cf}); an extensive computational
study supports the strong conjecture that this ratio is the actual
worst case across all Collatz orbits at sampled scales
(Conjecture~\ref{conj:65-41}); and the same arithmetic input
(transcendence of $\log_2 3$) yields a clean cycle-element upper
bound (Corollary~\ref{cor:cycle}).  None of these results resolves
the Collatz conjecture, but together they pinpoint the exact
arithmetic obstruction that any future quantitative theory will need
to address.

%% =================================================================
\begin{thebibliography}{99}
%% =================================================================

\bibitem{Baker1990}
A.~Baker,
\emph{Transcendental Number Theory},
2nd ed., Cambridge University Press, Cambridge, 1990.

\bibitem{BarinaCheckup}
D.~Barina,
\emph{Convergence verification of the Collatz problem},
J.\ Supercomput.\ \textbf{77} (2021), 2681--2688.

\bibitem{BernsteinLagarias1996}
D.\,J.~Bernstein and J.\,C.~Lagarias,
\emph{The $3x+1$ conjugacy map},
Canad.\ J.\ Math.\ \textbf{48} (1996), 1154--1169.

\bibitem{Bugeaud2004}
Y.~Bugeaud,
\emph{Approximation by Algebraic Numbers},
Cambridge Tracts in Mathematics, vol.~160, Cambridge University
Press, Cambridge, 2004.

\bibitem{ChangPaperA}
L.~Chang,
\emph{A three-state finite automaton for carry propagation in the
accelerated Collatz map: spectral gap and closed-form shift dynamics},
preprint, 2026.

\bibitem{ChangPaperB}
L.~Chang,
\emph{No finite local potential function is a strict Lyapunov
function for the accelerated Collatz map: an impossibility result},
preprint, 2026.

\bibitem{Conway1972}
J.\,H.~Conway,
\emph{Unpredictable iterations},
in: Proc.\ 1972 Number Theory Conference, Univ.\ of Colorado, 1972,
pp.\ 49--52.

\bibitem{Crandall1978}
R.\,E.~Crandall,
\emph{On the ``$3x+1$'' problem},
Math.\ Comp.\ \textbf{32} (1978), 1281--1292.

\bibitem{Eliahou1993}
S.~Eliahou,
\emph{The $3x+1$ problem: new lower bounds on nontrivial cycle lengths},
Discrete Math.\ \textbf{118} (1993), 45--56.

\bibitem{Gelfond1934}
A.~Gelfond,
\emph{Sur le septi\`eme probl\`eme de Hilbert},
Bull.\ Acad.\ Sci.\ URSS \textbf{7} (1934), 623--634.

\bibitem{HardyWright}
G.\,H.~Hardy and E.\,M.~Wright,
\emph{An Introduction to the Theory of Numbers},
6th ed., Oxford University Press, Oxford, 2008.

\bibitem{Khinchin1964}
A.\,Y.~Khinchin,
\emph{Continued Fractions},
University of Chicago Press, Chicago, 1964.

\bibitem{Lagarias1985}
J.\,C.~Lagarias,
\emph{The $3x+1$ problem and its generalizations},
Amer.\ Math.\ Monthly \textbf{92} (1985), 3--23.

\bibitem{Lagarias2010}
J.\,C.~Lagarias (ed.),
\emph{The Ultimate Challenge: The $3x+1$ Problem},
American Mathematical Society, Providence, RI, 2010.

\bibitem{Schneider1934}
T.~Schneider,
\emph{Transzendenzuntersuchungen periodischer Funktionen, I:
Transzendenz von Potenzen},
J.\ reine angew.\ Math.\ \textbf{172} (1934), 65--69.

\bibitem{Steiner1977}
R.\,P.~Steiner,
\emph{A theorem on the syracuse problem},
in: Proc.\ 7th Manitoba Conf.\ on Numerical Math.\ and Computing,
Congress.\ Numer.\ \textbf{20}, Utilitas Math., Winnipeg, 1977,
pp.\ 553--559.

\bibitem{Tao2022}
T.~Tao,
\emph{Almost all orbits of the Collatz map attain almost bounded
values},
Forum Math.\ Pi \textbf{10} (2022), e12.

\bibitem{Wirsching1998}
G.\,J.~Wirsching,
\emph{The Dynamical System Generated by the $3n+1$ Function},
Lecture Notes in Math.\ \textbf{1681}, Springer, Berlin, 1998.

\end{thebibliography}

\end{document}
